Big Ideas in Sports Analytics and Statistical Tools for their Investigation

### 2022-10-17 sports_analytics.Rmd
a couple of paragraphs on Elo (#36)

Another widely known tool for measuring team strength is the Elo rating system [@Elo1978], which was originally developed for chess, but also has been used to estimate team strength in other sports. 
See @koning2017rating for more information on applications of the Elo rating in soccer, and @kovalchik2019calibration and @kovalchik2016searching for Elo ratings in tennis. 
Elo ratings are used extensively for rankings of teams in numerous sports by the data journalists at [FiveThirtyEight.com](https://fivethirtyeight.com).

>>>>

<!-- a basic Elo probability model for estimating the probability of $A$ beating $B$ is -->
<!-- $$ -->
<!-- P(A) = W(R_A - R_B) \,. -->
<!-- $$ -->
<!-- Here, $W$ is a continuous and monotonically increasing function that maps the difference in strength to (0, 1).  -->

Another widely known tool for measuring team strength is the Elo rating system [@Elo1978], which was originally developed for chess.
Given two players $i$ and $j$ with unknown ratings $R_i$ and $R_j$, the probability $\Pi_{ij}$ of $i$ beating $j$ is defined as
$$
\Pi_{ij} = \frac{1}{1 + K^{(R_j - R_i)/400}} \,.
$$
In this formula, $K$ is commonly known as the $K$-factor (or development coefficient).
The International Chess Federation (FIDE) chooses $K=10$ for the calculation of Elo ratings.

Another interpretation for $\Pi_{ij}$ is the expected score of the game for player $i$.
The scores of 0, 0.5, and 1 are associated with three possible game outcomes loss, tie, and win, respectively.
After a game, the updated Elo rating $R^*_i$ for player $i$ is
$$
R^*_i = R_i + K(S_i - \Pi_{ij}) \,,
$$
where $S_i \in \{0, 0.5, 1\}$.
When a tournament concludes, a post-tournament rating is obtained for each player based on the rating updates for all games played.

[Mention glicko as an alternative to improve Elo]

Besides chess, the Elo system has also been implemented to estimate team strength in other sports. 
See @koning2017rating for more information on applications of the Elo rating in soccer, and @kovalchik2019calibration and @kovalchik2016searching for Elo ratings in tennis. 
Furthermore, Elo ratings are used extensively for rankings of teams in numerous sports by the data journalists at [FiveThirtyEight.com](https://fivethirtyeight.com).
